---
title: "tweet_shot"
author: "John Van Camp"
date: "12/31/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A Twitter discussion:

    I'm going to keep my eyes out for this one! Would love to have an easy way to embed tweets in Rmd talks!

    — Jeff Hollister (@jhollist) December 30, 2017

that spawned from Maëlle’s recent look-back post turned into a quick function for capturing an image of a Tweet/thread using webshot, rtweet, magick and glue.

Pass in a status id or a twitter URL and the function will grab an image of the mobile version of the tweet.

The ultimate goal is to make a function that builds a tweet using only R and magick. This will have to do until the new year.

```{r message=FALSE, warning=FALSE}

require(glue, quietly=TRUE)
require(rtweet, quietly=TRUE)
require(magick, quietly=TRUE)
require(webshot, quietly=TRUE)

tweet_shot <- function(statusid_or_url, zoom=3) {
  
  x <- statusid_or_url[1]

  is_url <- grepl("^http[s]://", x)

  if (is_url) {

    is_twitter <- grepl("twitter", x)
    stopifnot(is_twitter)

    is_status <- grepl("status", x)
    stopifnot(is_status)

    already_mobile <- grepl("://mobile\\.", x)
    if (!already_mobile) x <- sub("://twi", "://mobile.twi", x)

  } else {

    x <- rtweet::lookup_tweets(x)
    stopifnot(nrow(x) > 0)
    x <- glue_data(x, "https://mobile.twitter.com/{screen_name}/status/{status_id}")

  }

  tf <- tempfile(fileext = ".png")
  on.exit(unlink(tf), add=TRUE)

  webshot(url=x, file=tf, zoom=zoom)

  img <- image_read(tf)
  img <- image_trim(img)

  if (zoom > 1) img <- image_scale(img, scales::percent(1/zoom))

  img

}

```

Example of use:
```{r}
tweet_shot("https://twitter.com/jhollist/status/947082036019388416")
```
