---
title: "Packages after R Upgrade"
author: "John Van Camp"
date: "1/25/2017"
output: html_document
---
## Upgrade R for Mac & RStudio

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To save and reinstall packages from an old version of R to a new one.
In the old version type:
```{r}
tmp <- installed.packages()
installedpkgs = as.data.frame(installed.packages(.libPaths()[1]), stringsAsFactors = F)
save(installedpkgs, file="installed_old.rda")
```

Install the most recent version of R:
Download the most recent version of R from The Comprehensive R Archive Network (CRAN)

Install BioConductor packages using the biocLite.R installation script.
In an R command window, type the following:
```{r}
source("http://bioconductor.org/biocLite.R")
chooseBioCmirror()
biocLite()
# If you have other Bioconductor packages missing from the old installation:
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
for (i in 1:length(missing)) biocLite(missing[i])
```

Re-install the missing packages from CRAN:
```{r}
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
install.packages(missing)
update.packages()
```

If you find your X11 broken after the installation procedure (it happens every time to me, at least on Leopard) install the XQuartz App from here.

